---
title: "Inquiring Tweets Want to Know: #Edchat Supports for #RemoteTeaching during COVID-19"
author: "K. Bret Staudt Willet"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
usethis::use_git_ignore(c("*.csv", "*.rds", ".log"))

library(tidyverse)
library(anytime)
library(lubridate)
library(viridis)
library(quanteda)
library(topicmodels)
library(tidytext)
library(seriation)
```

## Look at all #Edchat tweets during COVID-19 (March-May 2020) vs. a year earlier (March-May 2019)

**Note:** These two .csv files are available upon request from https://osf.io/. Upon retrieval  from OSF, put the files in the same directory as the R markdown file you are currently reading. This file should be fully functional then.

```{r, include=FALSE, message=FALSE}
tweet_data_2020_0 <- 
  read_csv("tweets-2020-mar-may.csv", 
           col_types = cols(status_id = col_character(),
                            user_id = col_character()
           )
  )

tweet_data_2019_0 <- 
  read_csv("tweets-2019-mar-may.csv", 
           col_types = cols(status_id = col_character(),
                            user_id = col_character()
           )
  )
```

```{r, include=FALSE}
date_start_2020 <- 
  as_datetime("2020-03-01 05:00:00 UTC") %>% 
  ymd_hms() %>%
  with_tz(tzone = "US/Eastern")
date_end_2020 <- 
  as_datetime("2020-06-01 03:59:59 UTC") %>% 
  ymd_hms() %>%
  with_tz(tzone = "US/Eastern")
date_start_2019 <- 
  as_datetime("2019-03-01 05:00:00 UTC") %>% 
  ymd_hms() %>%
  with_tz(tzone = "US/Eastern")
date_end_2019 <- 
  as_datetime("2019-06-01 03:59:59 UTC") %>% 
  ymd_hms() %>%
  with_tz(tzone = "US/Eastern")
```

```{r, include=FALSE}
tweet_data_2020 <-  
  tweet_data_2020_0 %>%
  mutate(created_at = created_at %>% 
           as.numeric() %>% 
           anytime(asUTC = TRUE) %>% 
           as_datetime %>%
           ymd_hms() %>%
           with_tz(tzone = "US/Eastern"),
         year = 2020,
         has_question = ifelse(grepl("\\? ", text), TRUE, FALSE)
  ) %>%
  filter(created_at >= date_start_2020,
         created_at <= date_end_2020)
rm(tweet_data_2020_0)

tweet_data_2019 <-  
  tweet_data_2019_0 %>%
  mutate(created_at = created_at %>% 
           as.numeric() %>% 
           anytime(asUTC = TRUE) %>% 
           as_datetime %>%
           ymd_hms() %>%
           with_tz(tzone = "US/Eastern"),
         year = 2019,
         has_question = ifelse(grepl("\\? ", text), TRUE, FALSE)
  ) %>%
  filter(created_at >= date_start_2019,
         created_at <= date_end_2019)
rm(tweet_data_2019_0)
```

```{r, include=FALSE}
n_orig_2020 <- tweet_data_2020 %>% filter(!is_retweet) %>% nrow()
p_orig_2020 <- round(100 * n_orig_2020  / nrow(tweet_data_2020),  2)
n_questions_2020 <- tweet_data_2020 %>% filter(!is_retweet, has_question) %>% nrow()
p_questions_2020 <- round(100 * n_questions_2020  / n_orig_2020,  2)
n_questioners_2020 <-
  tweet_data_2020 %>% 
  filter(!is_retweet, has_question) %>% 
  pull(user_id) %>%
  unique() %>%
  length()

n_orig_2019 <- tweet_data_2019 %>% filter(!is_retweet) %>% nrow()
p_orig_2019 <- round(100 * n_orig_2019 / nrow(tweet_data_2019),  2)
n_questions_2019 <- tweet_data_2019 %>% filter(!is_retweet, has_question) %>% nrow()
p_questions_2019 <- round(100 * n_questions_2019 / n_orig_2019,  2)
n_questioners_2019 <-
  tweet_data_2019 %>% 
  filter(!is_retweet, has_question) %>% 
  pull(user_id) %>%
  unique() %>%
  length()
```

Of the **`r nrow(tweet_data_2020)` total tweets** in #Edchat between March 1 and May 31, 2020, there are **`r n_orig_2020` original tweets** (i.e., not retweets), or `r p_orig_2020`% of all tweets. Of the  `r n_orig_2020` original tweets, **`r n_questions_2020`** (`r p_questions_2020`%) contained a question.

Of the **`r nrow(tweet_data_2019)` total tweets** in #Edchat between March 1 and May 31, 2019, there are **`r n_orig_2019` original tweets** (i.e., not retweets), or `r p_orig_2019`% of all tweets. Of the  `r n_orig_2019` original tweets, **`r n_questions_2019`** (`r p_questions_2019`%) contained a question.

### Daily tweets

```{r, include=FALSE, message=FALSE}
to_plot_retweets_2020 <-
  tweet_data_2020 %>% 
  filter(is_retweet) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Retweet")

to_plot_orig_2020 <- 
  tweet_data_2020 %>%
  filter(!is_retweet,
         !has_question) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Original")

to_plot_questions_2020 <- 
  tweet_data_2020 %>%
  filter(!is_retweet,
         has_question) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Question")

to_plot_2020 <- 
  to_plot_retweets_2020 %>% 
  full_join(to_plot_orig_2020, by = c("day", "type", "n")) %>%
  full_join(to_plot_questions_2020, by = c("day", "type", "n"))

to_plot_retweets_2019 <-
  tweet_data_2019 %>% 
  filter(is_retweet) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Retweet")

to_plot_orig_2019 <- 
  tweet_data_2019 %>%
  filter(!is_retweet,
         !has_question) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Original")

to_plot_questions_2019 <- 
  tweet_data_2019 %>%
  filter(!is_retweet,
         has_question) %>%
  pull(created_at) %>% 
  floor_date("day") %>% 
  as_date() %>%
  table() %>% 
  as_tibble() %>%
  rename(day = ".") %>%
  mutate(day = as_date(day),
         type = "Question")

to_plot_2019 <- 
  to_plot_retweets_2019 %>% 
  full_join(to_plot_orig_2019, by = c("day", "type", "n")) %>%
  full_join(to_plot_questions_2019, by = c("day", "type", "n"))
```

```{r,  include=FALSE, message=FALSE}
to_plot_2020_adjusted <-
  to_plot_2020 %>%
  mutate(year = as.character(year(day)))
to_plot_2020_adjusted$day <- ymd(to_plot_2020_adjusted$day) - years(1) + days(2)

to_plot_2019_adjusted <-
  to_plot_2019 %>%
  mutate(year = as.character(year(day)))

to_plot_comparison <- bind_rows(to_plot_2019_adjusted, to_plot_2020_adjusted)
```

```{r tweets-over-time-comparison, include=TRUE, message=FALSE, echo=FALSE, fig.width=16, fig.height=9}
ggplot(data = to_plot_comparison, 
       mapping = aes(x = day, 
                     y = n, 
                     color = year,
                     shape = type)) +
  geom_point(size = 5, alpha= .7) +
  geom_line(size = 1, alpha= 1) +
  scale_color_manual(values = c("#63ACBE", "#EE442F")) + 
  scale_shape_manual(values=c(16, 18, 17)) +
  #geom_smooth(method='lm', se=FALSE, size=1) +
  xlab(NULL) +
  ylab("Number of Tweets") +
  theme_bw() +
  theme(panel.grid.major = element_line(color = "gray30"),
        panel.grid.minor = element_line(color = "gray90"),
        axis.title=element_text(size=28, family='serif'),
        axis.text=element_text(size=24, family='serif'),
        legend.position='bottom',
        legend.box = 'vertical',
        legend.box.background = element_rect(),
        legend.title=element_text(size=28, family='serif'), 
        legend.text=element_text(size=24, family='serif')
  ) +
  labs(color='Year', shape='Tweet Type')
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/fig1-tweets-over-time-comparison.png", width = 16, height = 9)
```

### Comparing top hashtags in 2020 vs. 2019

```{r, include=FALSE}
hashtag_table_2020 <-
  tweet_data_2020 %>%
  pull(hashtags) %>% 
  tolower() %>% 
  str_split(pattern = ", ") %>%
  unlist() %>%
  table() %>% 
  as_tibble() %>%
  rename(hashtag = ".",
         n_2020 = n) %>%
  filter(hashtag != "edchat") %>%
  mutate(p_2020 = round(100 * n_2020 / n_orig_2020, 2),
         odds_2020 = (n_2020 / n_orig_2020) / ((n_orig_2020 - n_2020) / n_orig_2020)
  ) %>%
  arrange(-odds_2020)

hashtag_table_2019 <-
  tweet_data_2019 %>%
  pull(hashtags) %>% 
  tolower() %>% 
  str_split(pattern = ", ") %>%
  unlist() %>%
  table() %>% 
  as_tibble() %>%
  rename(hashtag = ".",
         n_2019 = n) %>%
  filter(hashtag != "edchat") %>%
  mutate(p_2019 = round(100 * n_2019 / n_orig_2019, 2),
         odds_2019 = (n_2019 / n_orig_2019) / ((n_orig_2019 - n_2019) / n_orig_2019)
  ) %>%
  arrange(-odds_2019)

hashtag_comparison_table_2020 <-
  hashtag_table_2020 %>%
  head(100) %>%
  left_join(hashtag_table_2019, by = 'hashtag') %>%
  mutate(across(n_2020:odds_2019, coalesce, 0))

hashtag_comparison_table_2019 <-
  hashtag_table_2019 %>%
  head(100) %>%
  left_join(hashtag_table_2020, by = 'hashtag') %>%
  mutate(across(n_2019:odds_2020, coalesce, 0))

hashtag_comparison_table <-
  hashtag_comparison_table_2020 %>%
  full_join(hashtag_comparison_table_2019, 
            by = c('hashtag', 'n_2020', 'p_2020', 'odds_2020', 
                   'n_2019', 'p_2019', 'odds_2019') 
  ) %>%
  mutate(hashtag = paste0("#", hashtag),
         log_odds_ratio = log(odds_2020 / odds_2019)) %>%
  arrange(odds_2020)

hashtags_comparison_plot <- 
  hashtag_comparison_table %>%
  mutate(log_odds_ratio = abs(log_odds_ratio)) %>%
  filter(p_2020 >= 0.9 | p_2019 >= 0.9,
         log_odds_ratio >= 0.35) %>%
  arrange(-log_odds_ratio)
```

```{r, echo=FALSE, warning=FALSE, fig.width=16, fig.height=9}
ggplot(data = hashtags_comparison_plot, 
       mapping = aes(x = odds_2019, y = odds_2020)) +
  geom_point(alpha = 0.8, 
             size = 15,
             show.legend = TRUE,
             aes(color = abs(log_odds_ratio))) +
  scale_color_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
      ifelse(x < 4, 
             scales::rescale(x,
                             to = to,
                             from = c(min(x, na.rm = TRUE), 4)),
             1)
      }
      ) +
  geom_text(aes(label = hashtag), 
            check_overlap = TRUE,
            nudge_y = 0,
            nudge_x = 0.0075,
            size = 7,
            family = 'serif') +
  geom_abline(color = "red") +
  theme_bw() + 
  xlab("Odds (2019)") +
  ylab("Odds (2020)") +
  xlim(0, .14) +
  ylim(0, .22) +
  geom_hline(yintercept = 0, color = "black") + 
  geom_vline(xintercept = 0, color = "black") +
  theme(panel.border = element_rect(color = "gray80"),
        panel.grid.major = element_line(color = "gray30"),
        panel.grid.minor = element_line(color = "gray80"),
        axis.title=element_text(size=28, family='serif'),
        axis.text=element_text(size=24, family='serif')
  ) +
  labs(color = 'log odds ratio')
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/fig2-hashtag-comparison-scatter.png", width = 16, height = 9)
```

## Look at #Edchat question tweets

### Comparing words and topics in 2020 vs. 2019 (only question tweets)

```{r, include=FALSE}
corpus_2019q <- 
  tweet_data_2019 %>%
  filter(!is_retweet,
         has_question) %>%
  mutate(text = gsub("#", "", text)) %>%
  rename(doc_id = status_id) %>%
  select(doc_id, text) %>%
  corpus(text_field = 'text')
  

dfm_2019q <- 
  quanteda::dfm(corpus_2019q, 
      remove = c("edchat", "#edchat", stopwords("english")),
      remove_url = TRUE,
      remove_punct = TRUE,
      remove_symbols = TRUE,
      remove_numbers = TRUE)

corpus_2020q <- 
  tweet_data_2020 %>%
  filter(!is_retweet,
         has_question) %>%
  mutate(text = gsub("#", "", text)) %>%
  rename(doc_id = status_id) %>%
  select(doc_id, text) %>%
  corpus(text_field = 'text')

dfm_2020q <- 
  quanteda::dfm(corpus_2020q, 
      remove = c("edchat", "#edchat", stopwords("english")),
      remove_url = TRUE,
      remove_punct = TRUE,
      remove_symbols = TRUE,
      remove_numbers = TRUE)
```

#### 2019 words

```{r, echo=FALSE, fig.width=9, fig.height=9}
dfm_2019q %>%
  textplot_wordcloud(min_count = 10, 
                     random_order = FALSE, 
                     rotation = 0.25, 
                     color = viridis_pal()(8)
  )
```

#### 2020 words

```{r, echo=FALSE, fig.width=9, fig.height=9}
dfm_2020q %>%
  textplot_wordcloud(min_count = 10, 
                     random_order = FALSE, 
                     rotation = 0.25, 
                     color = viridis_pal()(8)
  )
```

#### 2019 topics

```{r, include=FALSE}
dfm_2019q <- dfm_2019q[rowSums(dfm_2019q) != 0, ]
lda_2019q <- dfm_2019q %>% LDA(k = 5, control = list(seed = 10272020))
topics_2019q <- tidytext::tidy(lda_2019q, matrix = "beta")

top_terms_2019q <- 
  topics_2019q %>%
  filter(term != "#edchat", term != "amp", term != "can") %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topic = str_c("2019-", topic))

#top_terms_2019$term <- 
#  factor(top_terms_2019$term, levels = sort(unique(top_terms_2019$term)))
```

```{r, echo=FALSE, warning=FALSE, fig.width=16, fig.height=9}
top_terms_2019q %>%
  #mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(x = term, 
             y  = beta, 
             fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = 'fixed') +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(top_terms_2019q$term))) +
  theme_bw() +
  xlab("Term") +
  ylab("Topic-Word Density") +
  theme(axis.title=element_text(size=28, family='serif'),
        axis.text=element_text(size=16, family='serif')
  )
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/topics-2019-questions.png", width = 16, height = 9)
```

#### 2020 topics

```{r, include=FALSE}
dfm_2020q <- dfm_2020q[rowSums(dfm_2020q) != 0, ]
lda_2020q <- dfm_2020q %>% LDA(k = 5, control = list(seed = 10272020))
topics_2020q <- tidytext::tidy(lda_2020q, matrix = "beta")

top_terms_2020q <- 
  topics_2020q %>%
  filter(term != "#edchat", term != "amp", term != "can") %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topic = str_c("2020-", topic))

#top_terms_2020$term <- 
#  factor(top_terms_2020$term, levels = sort(unique(top_terms_2020$term)))
```

```{r, echo=FALSE, warning=FALSE, fig.width=16, fig.height=9}
top_terms_2020q %>%
  #mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(x = term, 
             y  = beta, 
             fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = 'fixed') +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(top_terms_2020q$term))) +
  theme_bw() +
  xlab("Term") +
  ylab("Topic-Word Density") +
  theme(axis.title=element_text(size=28, family='serif'),
        axis.text=element_text(size=16, family='serif')
  )
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/topics-2020-questions.png", width = 16, height = 9)
```

```{r, include=FALSE}
top_terms_2019_2020q <-
  top_terms_2019q %>%
  full_join(top_terms_2020q, by = c("topic", "term", "beta")) %>%
  complete(topic, term, fill = list(beta = 0)) %>%
  filter(term != "education")

top_terms_2019_2020_matrixq <-
  top_terms_2019_2020q %>%
  pivot_wider(names_from = topic,
              values_from = beta) %>%
  column_to_rownames('term') %>%
  as.matrix()
  
top_terms_2019_2020_seriationq <-  
  top_terms_2019_2020_matrixq %>%
  seriation::seriate()

terms_orderedq <-
  unique(top_terms_2019_2020q$term)[get_order(top_terms_2019_2020_seriationq, dim=1)]
topics_orderedq <-
  unique(top_terms_2019_2020q$topic)[get_order(top_terms_2019_2020_seriationq, dim=2)]

top_terms_2019_2020_to_plotq <-
  top_terms_2019_2020q %>%
  mutate(topic = factor(topic, levels = unique(top_terms_2019_2020q$topic)),  # topics_orderedq
         term = factor(term, levels = terms_orderedq)
  )
```

```{r, echo=FALSE, fig.width=16, fig.height=12}
ggplot(top_terms_2019_2020_to_plotq, aes(x = topic, y = term)) +
        geom_tile(aes(fill = beta), color = 'white', size = 0.2) + 
        scale_fill_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
                ifelse(x < .012, 
                       scales::rescale(x,
                                       to = to,
                                       from = c(min(x, na.rm = TRUE), .012)), 
                       1)
        }
        ) +
        xlab("Topic") +
        scale_x_discrete(position='bottom') +
        ylab("Term") +
        scale_y_discrete(position='right') +
        theme_bw() +
        theme(axis.title = element_text(size=28, family='serif'),
              axis.text = element_text(size=20, family='serif'),
              legend.title = element_text(size=14, family='serif'), 
              legend.text = element_text(size=14, family='serif'),
              legend.position='left',
              legend.background = element_blank(),
              legend.box.background = element_rect(colour='black'),
              panel.background = element_blank(),
              panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              plot.background = element_blank()
        ) +
        labs(fill = "Topic-Word Density")
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/fig3-topics-2019-2020-questions.png", width = 16, height = 12)
```

### Look at high volume tweeters asking questions in 2020

```{r, include=TRUE}
question_tweets_2020 <-
  tweet_data_2020 %>%
  filter(!is_retweet,
         has_question) %>%
  mutate(screen_name = tolower(screen_name),
         month = lubridate::month(created_at),
         month = ifelse(month==3, 
                        "March", 
                        ifelse(month==4, 
                               "April", 
                               "May")
                        ),
         month = factor(month, levels = c("March", "April", "May"))
         )
questioner_table_2020 <-
  question_tweets_2020 %>%
  pull(screen_name) %>% 
  table() %>% 
  as_tibble() %>%
  arrange(desc(n)) %>%
  rename(screen_name = ".")
```

**`r nrow(questioner_table_2020)`** tweeters posted **`r n_questions_2020`** question tweets.

```{r, include=FALSE}
questioner_high_freq_2020 <-
  questioner_table_2020 %>%
  filter(n >= 10) %>%
  full_join(question_tweets_2020, by = 'screen_name') %>%
  filter(screen_name %in% questioner_high_freq_table_2020$screen_name)
```

**`r nrow(questioner_high_freq_table_2020)`** #Edchat participants contributed at least 10 question tweets during this time period (*n* = `r nrow(questioner_high_freq)` tweets from these high frequency question-askers). This means that **`r round(100 * nrow(questioner_high_freq_table_2020) / nrow(questioner_table_2020), 2)`%** of question-askers asked **`r round(100 * nrow(questioner_high_freq_2020) / n_questions_2020, 2)`%** of the questions.

```{r, echo=FALSE}
ggplot(data = questioner_high_freq_2020,
       aes(x = n)) +
  geom_boxplot(outlier.colour = 'black', 
               outlier.shape = 19,
               outlier.size = 2) +
  stat_boxplot(geom ='errorbar',
               coef = 1.50,
               width = 0.1) +
  theme_bw()
```

```{r, include=FALSE, eval=FALSE}
ggsave("output/questioner-number-of-questions-boxplot.png", width = 16, height = 9)
```

```{r session, include=TRUE}
devtools::session_info()
```
